Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-15 manuscript/draft.Rmd
stuff

Forecasting the future state of ecological systems is increasingly important for
planning and management, and also for quantitatively evaluating how well our
models capture the key ecological processes governing these systems [@clark2001;
@houlahan2017; @dietze2017]. Forecasts regarding biodiversity are especially
important, due to biodiversity's central role in conservation planning and its
sensitivity to anthropogenic effects [@cardinale2012; @diaz2015; @tilman2017].
High-profile studies forecasting large biodiversity declines over the coming
decades have played a large role in shaping ecologists' priorities [as well as
those of policymakers; @ipcc2014], but it is inherently difficult to evaluate
such long-term predictions before the projected biodiversity declines have
occurred.

@@

**Conclusions.** The science of forecasting biodiversity, and ecology more
broadly, remains in its infancy and it is important to consider the general
inability of forecasting methods to improve on simple baselines in that context.
When weather forecasting first started the forecasts were likewise worse than
simple baselines [@mcgill2012]. One of the things that helped weather forecasts
improve was large numbers of forecasts were made in public, which allowed
different approaches to forecasting to be regularly assessed and improved
[@mcgill2012; @silver2012]. This suggests that it is important for ecologists to
start regularly making and evaluating true ecological forecasts, even if they
perform poorly, and to make these forecasts publicly available for assessment.
These forecasts should include both short-term predictions, which can be
assessed quickly, and mid to long-term forecasts to help assess long time-scale
processes and determine how far into the future we can successfully forecast
[@tredennick2016; @dietze2017]. Forecasts based on the models in this paper from
now until 2050 are openly archived on Dryad so that we and others can assess how
well they perform and we plan to evaluate these forecasts and report the results
as each new year of BBS data becomes available. Weather forecasting has
continually improved throughout its history [@bauer2015] in part due to making
and evaluating public forecasts [@mcgill2012] and we hope this will help ecology
do the same.

Making successful ecological forecasts will be challenging. Ecological systems
are complex, fundamental theory is less refined than simpler physical/chemical
systems, and we currently lack the large amounts of data that produce effective
forecasts through machine learning. Despite this we believe that progress can be
made if we build an active forecasting culture in ecology that builds and
assesses forecasts in ways that will allow us to improve the effectiveness of
ecological forecasts most rapidly (Box 1). This includes expanding the scope of
the ecological and environmental data we work with, paying attention to
uncertainty in both model building and forecast evaluation, and rigorously
assessing forecasts using a combination of hindcasting, archived forecasts, and
comparisons to simple baselines.

@@

Typically ecological forecasts use one modeling approach or a small number of
related approaches. By fitting and evaluating multiple modeling approaches we
can learn more rapidly about the best approaches for making predictions for a
given ecological quantity. This includes comparing process based and data-driven
models and comparing the accuracy of forecasts to simple baselines to determine
if the modeled forecasts are more accurate than the naive assumption that the
world is static [@ye2015].

@@

Forecasts describe how systems are expected to change through time. While some
areas of ecological forecasting focus primarily on time-series data, others
primarily focus on using spatial models and space-for time substitutions. Using
ecological and environmental time-series data allows the consideration of actual
dynamics from both a process and error structure perspective [@tredennick2016].

@@

Understanding uncertainty in a forecast is just as important as understanding
the average or expected outcome. Failing to account for uncertainty can result
in overconfidence in highly uncertain outcomes leading to poor decision making
and erosion of confidence in ecological forecasts. Models should explicitly
include sources of uncertainty and propagate them through the forecast where
possible [@clark2001; @dietze2017]. Evaluations of forecasts should assess the
accuracy of uncertainties as well as point estimates [@dietze2017].

@@

In general the accuracy of forecasts decreases with the length of time into the
future being forecast [@petchey2015]. This decay in accuracy and the potential for
different rates of decay to result in different relative model performance at
different lead times should be considered when evaluating forecasts and
comparing models.

@@

Ecological observations are influenced by both the underlying processes and how
the system is sampled. When possible forecasts should model the factors
influencing the observation of the data. [[citation?]]

@@

Evalutating a models predictive performance across time is critical. Hindcasting
uses a temporal out of sample validation to mimic how well a model would have
performed had it been run in the past. For example using occurance data from the
early 20th century to model distributions which are validated with late
20th century occurances. Dense time series, such as yearly observations, are
desirable to also evalulate the forecast horizon (see #6), but this is not
a strict requirement.

@@

To allow the future evaluation of the accuracy and uncertainty of forecasts the
forecast values and/or models should be archived so that they can be assessed
after new data is generated [@mcgill2012; @silver2012]. Enough information
should be provided to allow an unambiguous assessment of the forecast
performance.

### 10. Make short-term and long-term predictions

In cases where long-term predictions are the primary goal, short-term should
also be made to accommodate the time-scales of planning and management decisions
and to allow the accuracy of the forecasts to be quickly evaluated
[@tredennick2016].

>>>>

Forecasting the future state of ecological systems is increasingly important for
planning and management, and also for quantitatively evaluating how well our
models capture the key ecological processes governing these systems [@clark2001;
@houlahan2017; @dietze2017]. Forecasts regarding biodiversity are especially
important, due to biodiversity's central role in conservation planning and its
sensitivity to anthropogenic effects [@cardinale2012; @diaz2015; @tilman2017].
High-profile studies forecasting large biodiversity declines over the coming
decades have played a large role in shaping ecologists' priorities [as well as
those of policymakers; e.g. @ipcc2014], but it is inherently difficult to
evaluate such long-term predictions before the projected biodiversity declines
have occurred.

@@

**Conclusions.** The science of forecasting biodiversity, and ecology more
broadly, remains in its infancy and it is important to consider the general
inability of forecasting methods to improve on simple baselines in that context.
When weather forecasting first started the forecasts were likewise worse than
simple baselines [@mcgill2012]. One practice that helped weather forecasts to
improve was publicly making a large number of predictions, which allowed
different forecasting approaches to be regularly assessed and refined
[@mcgill2012; @silver2012]. This suggests that it is important for ecologists to
start regularly making and evaluating ecological forecasts on real data, even if
they perform poorly, and to make these forecasts publicly available for
assessment. These forecasts should include both short-term predictions, which
can be assessed quickly, and mid- to long-term forecasts, which can help
ecologists to assess long time-scale processes and determine how far into the
future we can successfully forecast [@tredennick2016; @dietze2017]. Forecasts
based on the models in this paper (from now until 2050) are openly archived on
Zenodo, so that we and others can assess how well they perform. We also plan to
evaluate these forecasts and report the results as each new year of BBS data
becomes available at the same URL [@zenodo]. Weather forecasting has continually
improved throughout its history [@bauer2015], in part due to making and
evaluating public forecasts [@mcgill2012], and we hope this practice will help
ecology do the same.

Making successful ecological forecasts will be challenging. Ecological systems
are complex, fundamental theory is less refined than simpler physical/chemical
systems, and we currently lack the large amounts of data that produce effective
forecasts through machine learning. Despite this, we believe that progress can be
made if we build an active forecasting culture in ecology that builds and
assesses forecasts in ways that will allow us to improve the effectiveness of
ecological forecasts most rapidly (Box 1). This includes expanding the scope of
the ecological and environmental data we work with, paying attention to
uncertainty in both model building and forecast evaluation, and rigorously
assessing forecasts using a combination of hindcasting, archived forecasts, and
comparisons to simple baselines.

@@

Typically ecological forecasts use one modeling approach or a small number of
related approaches. By fitting and evaluating multiple modeling approaches we
can learn more rapidly about the best approaches for making predictions for a
given ecological quantity. This includes comparing process-based [e.g. [[refs]]]
and data-driven models, as well as comparing the accuracy of forecasts to simple
baselines to determine if the modeled forecasts are more accurate than the naive
assumption that the world is static [@ye2015].

@@

Forecasts describe how systems are expected to change through time. While some
areas of ecological forecasting focus primarily on time-series data, others
primarily focus on using spatial models and space-for-time substitutions. Using
ecological and environmental time-series data allows the consideration of actual
dynamics from both a process and error structure perspective [@tredennick2016].

@@

Understanding uncertainty in a forecast is just as important as understanding
the average or expected outcome. Failing to account for uncertainty can result
in overconfidence in highly uncertain outcomes leading to poor decision making
and erosion of confidence in ecological forecasts. Models should explicitly
include sources of uncertainty and propagate them through the forecast where
possible [@clark2001; @dietze2017]. Evaluations of forecasts should assess the
accuracy of models' estimated uncertainties as well as their point estimates
[@dietze2017].

@@

In general, the accuracy of forecasts decreases with the length of time into the
future being forecast [@petchey2015]. This decay in accuracy (and the potential
for different rates of decay to result in different relative model performance
at different lead times) should be considered when evaluating forecasts
and comparing models.

@@

Ecological observations are influenced by both the underlying biological
processes (e.g. resource limitation) and how the system is sampled. When
possible, forecasts should model the factors influencing the observation of the
data. [[citation?]]

@@

Evalutating a model's predictive performance across time is critical.
Hindcasting uses a temporal out-of-sample validation approach to mimic how well
a model would have performed had it been run in the past. For example, using
occurance data from the early 20th century to model distributions which are
validated with late 20th century occurances. Dense time series, such as yearly
observations, are desirable to also evalulate the forecast horizon (see #6), but
this is not a strict requirement.

@@

Forecast values and/or models should be archived so that they can be assessed
after new data is generated [@mcgill2012; @silver2012]. Enough information
should be provided in the archive to allow ecologists to make an unambiguous
assessment of each forecast's performance.

### 10. Make both short-term and long-term predictions

In cases where long-term predictions are the primary goal, short-term
predictions should also be made to accommodate the time-scales of planning and
management decisions and to allow the accuracy of the forecasts to be quickly
evaluated [@tredennick2016].
